\TheTitle

### 2016-04-25 linvis.tex
manymore changes

\newcommand{\TheTitle}{Visualizing Release Information of Linux}
\newcommand{\TheAuthors}{Evan Wilde, Daniel German}
\newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
\newcommand{\TheSubject}{Digesting large amounts of commit data}
\newcommand{\TheKeywords}{Linux, git, data structures, tree data structures}

@@

        In this paper, we present a tool that uses a tree-based model, as described
        by German et al., to construct a clearer summary of the changes made. This
        tool is designed to allow users to explore the topology of the repository
        for a given release of the kernel and view an aggregated summary of the
        changes made.

@@

This representation works in smaller projects; it enables users to see
when changes are made, when these changes are merged, and how each branch is
interacting.  
In large modular projects, like the Linux kernel, the DAG becomes a
mess of merges and commits (figure \ref{fig:gitk}) losing its visual meaning.  In
some cases, the Linux kernel is simply too large for the system to generate a
visualization (eg. Github provides a DAG view for many projects, but is unable to
display the visualization for projects as large as the Linux kernel (figure
\ref{fig:gitfail}).
Furthermore, Linux creates between 60 and 70k new commits every year and, according to our previous work \cite{German2015}pdfl, 
a commit takes a median of 30 days (from the moment it is authored) to the time it arrives to the official repository
This means that, if a person inspected the commits in this repository today, tomorrow it might contain  many more
of commits that were authored long before. However, distinguishing these new commits is not trivial.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/gitk.png}
        \caption{A sample DAG view of commits from Linux in Gitk}
        \label{fig:gitk}
\end{figure}

\begin{figure}
        \centering
        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
        \caption{Github Failure Message showing the DAG of the Linux Project}
        \label{fig:gitfail}
\end{figure}

One of the challenges with visualizing the arrival of commits to a repository is that
in Git, commits are immutable, and Git does not store the date that a commit was merged into another branch (including master).
To complicate the problem, the DAG points to the ancestors of a commit (a model necessary for the operation of Git), but developers 
would prefer to know the path that a commit followed to reach the master repository. \dmg{perhaps a figure is good here}.
Hence, tracing the path that any commit followed to master would imply that for any given merge, it would be possible to know what commits were merged. A user could
inspect the commits that have arrived into master within a given time frame by simply checking which commits where merged during that time frame.

@@

 The contributions of this paper are two. First, we describe a method to convert the DAG of the Linux repository into a tree (we call this the \emph{tree-model of the
 repository}) that represents the path used by
 any commit to reach master. Second, 
 we present a method to inspect and visualize the history of merges in the  Linux project using the tree-model of the repository

These methods and visualizations have been implemented in a web-based
tool\footnote{Our visualization tool is currently available at \url{http://li.turingmachine.org}}
instead of the DAG. Our visualizations and tool provide information about the location of any given
commit in the respective merge tree, the files edited, the modules edited, and the
commit message for a given commit or merge. Our tool allows users to apply various
filters, including the release version, along with a keyword or phrase from the log
preview, the name of the author, or the commit ID. The user can request all the
root-level merges containing a commit or merge that matches the query, or all
commits and merges that match the query.

\evantodo{move to discussion}

Using the tree provides many advantages to the DAG model, but there are two
drawbacks. The first drawback is that a server must query the state of the master
repository since time information is not stored by git. The second drawback is that
the tree construction is a heuristic approach, and in some cases may break.

Our core model provides advantages to the DAG model used by other tools. We separate
the commits and merges into sub-trees and focus on a single sub-tree at a time. This
enables us to filter out commits and merges that do not pertain to the component of
the kernel that we are interested in, providing a clearer picture of what is
happening in a given merge. The tree model also enables us to aggregate metadata
easily. With the addition of the continuous polling of the repository, the model
provides us with a mechanism of performing date-range queries with meaningful
results.

\section{Merge-Tree model}

@@

Computing the tree-model of a DAG of any repository might not be possible. However, certain features of the development process of Linux make it feasible to
compute the its merge-tree model. In particular, the Master branch of Linux is maintained by Linus Torvalds and only him has write access to it. We have verified this
assertion if previous research~\cite{German2015}. We have developed a heuristic that is presented in Algorithm~\ref{fig:alg}. In a nutshell, first identify the commits directly made
to the master branch. Then,
recursively, for
every commit, using the DAG, determine the shortest path (in terms of time) from such commit to master using the inverted DAG.

\begin{algorithm}
\caption{Computing the merge-tree of Linux Git's DAG}\label{fig:alg}
\begin{algorithmic}
\Function{ComputeTreeModel}{DAG}: tree 
\State {\# Compute the tree from the DAG of a git repository.}
\State {\#  Returns $Tree$, a graph containing every commit }
\State {\# in DAG with the path it followed to master.\\}

\State $head \gets \textit{Head of master of git repository}$
\State $master \gets \textit{traverse DAG from head using }$
\State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
\State $nodes(Tree) \gets nodes(DAG)$
\State \Function{distance2Master}{cid} : seconds 
\State {\# Compute the shortest distance to master setting}
\State {\# cid's successor (next) in its way to master.}
\State {\# This function should be memoized. Otherwise it}
\State {\# would run in exponential time.}
\If {\textit{cid in master}} 
\State \Return 0
\EndIf
\State    $d \gets 	\infty$
   \State {\# Traverse the inverted DAG}
   \For{$c \in children(cid, DAG)$} 
      \If {$c \in master$} 
      \State $d_1 \gets commitTime(c)-commitTime(cid)$
      \Else
      \State {$d_1 \gets distance2Master(c)$} 
      \EndIf
    \If {$d_1 < d $} 
    \State $next \gets c$
   \State  $d \gets d_1$
   \EndIf
   \EndFor
   \State {\# $c$ is the commit that follows $cid$}
   \State {\# in its way to master}
   \State add edge $(cid, next)$ to $Tree$
   \State \Return $d$ 
\EndFunction

@@

@@

The goal is to build a tool that simplifies  navigation of the kernel commit information, while providing an explanation of the changes occurring at each commit
and, more importantly, indicates how commits are merged in their way to master.
We designed the tool with two use-cases in mind, though a user may switch
between the cases as they work.

\noindent \textbf{Use-case 1: top-to-bottom approach}\label{sec:usecase1}\\
These are users that are maintaining a section of the kernel and would like to pick
a merge (including all the commits that it merges) and merge it directly into their current repository. This is useful for
reducing the amount of re-implementation work. For these users, it is important to have the
ability to aggregate metadata about files and modules being effected by the merge.
\dmg{what is root?, what is leaves? perhaps remove unless you can be specific}
It is also important for these users to be able to navigate from the root toward the
leaves.

\noindent \textbf{Use-case 2: bottom-to-top approach}\label{sec:usecase2}\\
These are users that start with a known merge or commit and would like to see what
other changes are being made in commits that are in the same merge. This is useful
to see what other commits are related to the current commit and how they get collated into merges that eventually end in master.
 This is primarily for maintainers that need to perform
some specific cherry picking of commits. We
must provide these users a mechanism for navigating from a single commit toward the
master, allowing them to see the bigger picture.

\subsection{Core Model}

In our visualizations, each commit is either made directly to Master, or is part of a Merge-tree, with its root in Master (a merge made by Linus Torvalds). Each merge may contain zero or more
children, which are other merges and commits.

To improve the visualization of the merge-tree, we transform the merge-tree into a new tree, where each commit points to its next merge in the way to Master. In
this model, commits have no children and are always leaf nodes. This allows the clustering of commits along their next merge, and avoid the space needed to
present long sequences of commits without merges in between.

Commits contain the
metadata for the changes made. This metadata includes the files changed, the lines
added and removed from each file, the author, the date the commit was merged into
the merge that led to being merged into the kernel, and the date the commit was
authored, the patch, and the commit log. Merges contain less metadata, only storing
the author of the merge, the log, the commit date, and the author date. The details
of the model are outlined in \cite{German2015}.

\section{Design}

\dmg{I would remove the command line shell, and use the space for related work}

We came up with two designs. We recognized that the structure of the repository was
similar to the wide and shallow tree-structure of a file system. We began with a
command-line shell interface. Then transitioned to a web-based tool where we looked
beyond the file system structure. In this section, we describe the progression of
our tool design, showing the benefits and drawbacks of each component.

\subsection{Command-line Shell}

Our initial intuition was to work with the commits and merges as directory trees
where the commits are files and the merges are directories. There are many
similarities between the tree structure of a file system and the tree structure of
the repository. Both trees are generally shallow, but very wide. File systems have
been studied for quite some time and there are many tools capable of visualizing and
working with the file structure.

The first tool we implemented was a small command-line shell (figure
\ref{fig:shell}) that allows users to navigate through the commits and merges as one
might navigate through files and directories. A user could navigate toward the
leaves of the tree by using the \verb|cd| command followed by the commit ID of the
next merge. The user could then navigate toward the root by issuing \verb|cd ../|
as they would in a file system. Users can use \verb|ls| to view what commits and
merges exist in the current working merge, then use \verb|cat| to view the log
preview of a commit or \verb|more| to see the preview and how many lines were added
and removed from the files in the commit.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/shell.png}
        \caption{Shell Design}
        \label{fig:shell}
\end{figure}

The shell does not meet our design goals. It only provides users with the ability to
navigate from the root. If there is a commit or merge that they would like to
investigate, they must find it with another tool and navigate through the tree to
find it. Furthermore, it does not aggregate information in the merges, so users must
manually aggregate the metadata. Another issue with this design is that it presents
the users with the commit hashes, which are quite long and unpleasant to remember
and type into the shell, but also do not provide any quick explanation of what the
commit or merge may contain.

\subsection{Web-based Tool}

We replace the command-line interface with a web-based tool. Creating a web-based
tool enables users to use the system without having to install additional software
or store a large database, making it more accessible, more easily maintainable, and
platform independent. A web-based tool can utilize more interactive means of
navigation and provide better explanations of commits and merges than the
command-line interface.

Our tool uses the following mechanisms to reach our goals of better navigation and
better explanation of the changes.

\begin{itemize}
        \item Filter by search
        \item View files edited
        \item View modules
        \item Tree viewer
\end{itemize}

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/search.png}
        \caption{Search View}
        \label{fig:search}
\end{figure}

Searching (figure \ref{fig:search}) allows a user to filter commits and merges that
are irrelevant. The search mechanism breaks down the results by release version. A
user can further narrow down the search range using the commit date range. The
commit date refers to the date of the creation of the top-level merge. The commits
and merges returned may have a commit date beyond the queried dates; the date-range
query refers to the date that the commit is finally merged into the kernel by Linus.
A commit may not be merged immediately, and in some cases may take years before
being applied. A user may then provide a search text, filtering by the author name,
the commit ID, or keyword from the log. Any part of the author name may show up in
the results, including searching by email address. The commit ID must be specified
from the first character to the last character. For example, the commit
`c267548755a184ef97301071300c1739a564e135' is returned when the user searches for a
commit ID of `c26' in the 3.10 Linux kernel. We will use this commit for showing
each of the features of the tool.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/search_results.png}
        \caption{Search Results}
        \label{fig:results}
\end{figure}

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/message_view.png}
        \caption{Commit Message}
        \label{fig:message}
\end{figure}

Once a user selects a commit or merge to investigate, they are presented with a
tabbed pane allowing them to view the full commit log, the files edited, the modules
involved, and the tree views.

The first tab displays the full commit log (figure \ref{fig:message}). From this, a
user is able to see what they would see had they searched for the commit using git
log. This doesn't provide additional information to the other tools, but helps to
complete the functionality of our tool. The commit log provides a user with the
information about the content of the commit and who has signed-off on the commit to
ensure that it is good quality. The message for merges may contain a summary of the
commits being merged.  The information within these messages is highly variable, and
is completely dependent on the author's style. As the user moves toward the
top-level merge, the quality of these messages generally improves.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/file_view.png}
        \caption{Commit Files}
        \label{fig:files}
\end{figure}

The second tab is the files tab (figure \ref{fig:files}). This tab provides
information on what files have been edited, how many lines were added, and how many
lines were removed in a given commit. This functionality is similar to the other
tools available. Our tree-based design model allows us to extend this functionality
to merges, which the other tools are unable to show. To find the number of lines
added to a file in a merge, we take the sum of the lines added to that file in each
of the children of that merge. We do the same for calculating the number of lines
removed. This could be extended to use the patch information to determine when a
line has been replaced rather than incrementing the lines added and removed.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/modules.png}
        \caption{Commit Modules}
        \label{fig:modules}
\end{figure}

The modules tab (figure \ref{fig:modules}) shows the modules that are contained
within the commit. Modules are not natively recognized by git, and are not going to
be present in all repositories.  In the Linux repository, authors put the module
they are working on in the one-line preview of the log-message. We heuristically
extract the module by taking all text in the log preview of commits until the first
colon.  Modules are logical partitions of the information in the kernel. Depending
on where the author was working, modules can be general, such as ``bluetooth'' and
``wireless'', or can be quite specific for individual hardware, such as
``ath9k\_hw'' and ``wl1251''. In a few cases, the author of a commit does not
correctly follow this format and the hueristic approach fails. Like files, we are
able to show what modules are edited in merges and commits. The image in figure
\ref{fig:modules} shows the modules for a commit.  As commits have no children, they
will only have a single module; merges with more than one child will have more.

Finally, we have the tree view tab. The tree view is designed for providing easy
navigation within a tree and providing a clear topological view of the branch of the
main kernel. We have experimented with various tree designs to find a design that
allows for easy navigation and visualization of both large and small trees. We
discuss the various tree designs in section \ref{treeview_section}.

\subsection{Tree Views} \label{treeview_section}

>>>>

\newcommand{\TheTitle}{Merge-tree: Visualizing the integration of commits into Linux}
\newcommand{\TheAuthors}{Evan Wilde, Daniel German}
\newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
\newcommand{\TheSubject}{Digesting large amounts of commit data}
\newcommand{\TheKeywords}{Linux, git, data structures, tree data structures}

@@

        In this paper, we present \tool, a tool that uses a tree-based model to construct a clearer summary of the changes made. \tool is designed to allow
        users to explore the topology of the repository for a given release of the kernel and view an aggregated summary of the changes made.

@@

This representation works in smaller projects; it enables users to see
when changes are made, when these changes are merged, and how each branch is
interacting.  
In large modular projects, like the Linux kernel, the DAG becomes a
mess of merges and commits (figure \ref{fig:gitk}) losing its visual meaning.  In
some cases, the Linux kernel is simply too large for the system to generate a
visualization (eg. Github provides a DAG view for many projects, but is unable to
display the visualization for projects as large as the Linux kernel (figure
\ref{fig:gitfail}).
Furthermore, Linux creates between 60 and 70k new commits every year and, according to our previous work \cite{German2015}, 
a commit takes a median of 30 days from the moment it is authored to the time it arrives to the official repository
This means that, if a person inspected the commits in this repository today, tomorrow it might contain  many more
commits that were authored long before. However, distinguishing these new commits is not trivial.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/gitk.png}
        \caption{A sample DAG view of commits from Linux in gitk. Commits point towards their ancestors, and there is no clear indication of the way the travel
          to reach Master. In fact, neither gitk nor git is capable of showing what commits were done in master.}
        \label{fig:gitk}
\end{figure}

\begin{figure}
        \centering
        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
        \caption{Github Failure Message showing the DAG of the Linux Project and its relationship to other forks.}
        \label{fig:gitfail}
\end{figure}

One of the challenges with visualizing the arrival of commits to a repository is that
in Git, commits are immutable, and Git does not store the date that a commit was merged into another branch (including master).
To complicate the problem, the DAG points to the ancestors of a commit (a model necessary for the operation of Git), but developers 
would prefer to know the path that a commit followed to reach the master repository.
Hence, tracing the path that any commit followed to master would imply that for any given merge, it would be possible to know what commits were merged. A user could
inspect the commits that have arrived into master within a given time frame by simply checking which commits where merged during that time frame.

@@

 The contributions of this paper are two. First, we describe a method to convert the DAG of the Linux repository into a tree (we call this the \emph{merge-tree of the
 repository}) that represents the path used by
 any commit to reach master. Second, 
 we present a method to inspect and visualize the history of merges in the  Linux project using the tree-model of the repository

These methods and visualizations have been implemented in a web-based
tool called \tool\footnote{\tool is currently available at \url{http://li.turingmachine.org}}. Our visualizations and tool provide information about the location of any given
commit in its respective merge-tree, the files edited, the modules edited, and the
commit message for a given commit or merge. \tool allows users to apply various
filters, including the release version, along with a keyword or phrase from the log
preview, the name of the author, or the commit ID. The user can request all the
root-level merges containing a commit or merge that matches the query, or all
commits and merges that match the query.

\section{Merge-Tree model}
\label{sec:mergetree}

@@

Computing the tree-model of a DAG of any repository might not be possible. However, certain features of the development process of Linux make it feasible to
compute its merge-tree. In particular, the Master branch of Linux is maintained by Linus Torvalds and only him has write access to it. We have verified this
assertion if previous research~\cite{German2015}. We have developed a heuristic that is presented in Algorithm~\ref{fig:alg}. In a nutshell, first identify the commits directly made
to the master branch. Then,
recursively, for
every commit, using the DAG, determine the shortest path (in terms of time) from such commit to master using the inverted DAG.

\begin{algorithm}
\caption{Computing the merge-tree of Linux Git's DAG}\label{fig:alg}
\begin{algorithmic}
\Function{ComputeMergeTree}{DAG}: tree 
\State {\# Compute the tree from the DAG of Linux repository.}
\State {\#  Returns $Tree$, a graph containing every commit }
\State {\# in DAG with the path it followed to master.\\}

\State $head \gets \textit{Head of master of git repository}$
\State $master \gets \textit{traverse DAG from head using }$
\State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
\State $nodes(Tree) \gets nodes(DAG)$
\State \Function{distance2Master}{cid} : seconds 
\State {\# Helper function}
\State {\# Recursively compute shortest distance to master}
\State {\#  setting cid's successor (next) in its way to master.}
\State {\# This function should be memoized. Otherwise it}
\State {\# would run in exponential time.}
\If {\textit{cid in master}} 
\State \Return 0
\EndIf
\State    $d \gets 	\infty$
   \State {\# Traverse the inverted DAG}
   \For{$c \in children(cid, DAG)$} 
      \If {$c \in master$} 
      \State $d_1 \gets commitTime(c)-commitTime(cid)$
      \Else
      \State {$d_1 \gets distance2Master(c)$} 
      \EndIf
    \If {$d_1 < d $} 
    \State $next \gets c$
   \State  $d \gets d_1$
   \EndIf
   \EndFor
   \State {\# $c$ is the commit that follows $cid$}
   \State {\# in its way to master}
   \State add edge $(cid, next)$ to $Tree$
   \State \Return $d$ 
\EndFunction

@@

The goal of \tool was to simplify the  navigation of the kernel commit information, specifically by focusing on merges.
This is done by leveraging the merge-tree view of Linux to inspect how commits are merged in their way to master.

\subsection{Use cases}

We designed \tool with two use-cases in mind, though a user may switch
between the cases as they work.

\noindent \textbf{Use-case 1: top-to-bottom approach}\label{sec:usecase1}\\
These are users that are maintaining a section of the kernel and would like to pick
a merge (including all the commits that it merges) and merge it directly into their current repository. This is useful for
reducing the amount of re-implementation work. For these users, it is important to have the
ability to aggregate metadata about files and modules being effected by the merge.
It is also important for these users to be able to navigate from the root of the merge-tree toward the
leaves.

\noindent \textbf{Use-case 2: bottom-to-top approach}\label{sec:usecase2}\\
These are users that start with a known merge or commit and would like to see what
other changes are being made in commits that are in the same merge, including knowing the merge-tree they belong to. This is useful
to see what other commits are related to the current commit and how they get collated into merges that eventually end in master.
 This is primarily for maintainers that need to perform
some specific cherry picking of commits. We
must provide these users a mechanism for navigating from a single commit toward the
master, allowing them to see other commits that might be related to it.

\subsection{Data Model}

In our visualizations, we leverage the Merge-Tree model described in Section~\ref{sec:mergetree}. In this model each commit is either already in Master, or is part of a tree which is rooted in the merge that merged it into Master. In that case, each
commit (whether merge or non-merge) has only one successor at most (the root of each tree has none, as it was made by Linus Torvalds directly into Master).
Commits contain the
metadata for the changes made. This metadata includes the files changed, the lines
added and removed from each file, the author, the date the commit was merged into
the merge that led to being merged into the kernel, and the date the commit was
authored, the patch, and the commit log. Merges contain less metadata, only storing
the author of the merge, the log, the commit date, and the author date, and potentially, changes necessary to address conflicts during the merge. The details
of the model are outlined in \cite{German2015}.

\section{Design and Implementation}

% \dmg{I would remove the command line shell, and use the space for related work}

% We came up with two designs. We recognized that the structure of the repository was
% similar to the wide and shallow tree-structure of a file system. We began with a
% command-line shell interface. Then transitioned to a web-based tool where we looked
% beyond the file system structure. In this section, we describe the progression of
% our tool design, showing the benefits and drawbacks of each component.

% \subsection{Command-line Shell}

% Our initial intuition was to work with the commits and merges as directory trees
% where the commits are files and the merges are directories. There are many
% similarities between the tree structure of a file system and the tree structure of
% the repository. Both trees are generally shallow, but very wide. File systems have
% been studied for quite some time and there are many tools capable of visualizing and
% working with the file structure.

% The first tool we implemented was a small command-line shell (figure
% \ref{fig:shell}) that allows users to navigate through the commits and merges as one
% might navigate through files and directories. A user could navigate toward the
% leaves of the tree by using the \verb|cd| command followed by the commit ID of the
% next merge. The user could then navigate toward the root by issuing \verb|cd ../|
% as they would in a file system. Users can use \verb|ls| to view what commits and
% merges exist in the current working merge, then use \verb|cat| to view the log
% preview of a commit or \verb|more| to see the preview and how many lines were added
% and removed from the files in the commit.

% \begin{figure}
%         \centering
%         \includegraphics[width=0.47\textwidth]{figures/shell.png}
%         \caption{Shell Design}
%         \label{fig:shell}
% \end{figure}

% The shell does not meet our design goals. It only provides users with the ability to
% navigate from the root. If there is a commit or merge that they would like to
% investigate, they must find it with another tool and navigate through the tree to
% find it. Furthermore, it does not aggregate information in the merges, so users must
% manually aggregate the metadata. Another issue with this design is that it presents
% the users with the commit hashes, which are quite long and unpleasant to remember
% and type into the shell, but also do not provide any quick explanation of what the
% commit or merge may contain.

% \subsection{Web-based Tool}

To navigate and inspect the merge-tree view of the Kernel we
created a web-based tool called \tool. Creating a web-based
tool enables users to use the system without having to install additional software
or store a large database, making it more accessible, more easily maintainable, and
platform independent. 
% A web-based tool can leverage interactive means of
% navigation and provide better explanations of commits and merges than the
% command-line interface.
\tool uses the following mechanisms to reach our goals of better navigation and
better explanation of the changes.

\begin{itemize}
        \item Filter by searching commits by when they are merged into Master.
        \item View files edited
        \item View modules
        \item Tree viewer
\end{itemize}

\subsection{Searching by when they are merged into Master}

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/search.png}
        \caption{Search View. It allows finding commits that were \textbf{merged} in a given period or by commit it.}
        \label{fig:search}
\end{figure}

Searching (depicted in Figure \ref{fig:search}) allows a user to filter commits and merges that are irrelevant. The search mechanism breaks down the results by
release version.  A user can further narrow down the search by \textbf{specifying a range of dates} in which such commits \textbf{were merged by Linus into
  Master, not
  when the commits were created.} This distinction is important. We have observed commits that take years
to arrive into Master after they were originally committed.

A user may then provide a search text, filtering by the author name,
the commit ID, or keyword from the log. Any part of the author name may show up in
the results, including searching by email address.

\dmg{What is this commit? any commit id, or a merge into Master? Unclear.}
If providing commit ID must be specified
from the first character to the last character. For example, the commit
`c267548755a184ef97301071300c1739a564e135' is returned when the user searches for a
commit ID of `c26' in the 3.10 Linux kernel. We will use this commit for showing
each of the features of \tool.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/search_results_2.png}
        \caption{Search Results. Each is a commit merged in the desired merged window.}
        \label{fig:results}
\end{figure}

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/log_view.png}
        \caption{Panel showing the Commit Message of a commit.}
        \label{fig:message}
\end{figure}

Once a user selects a commit or merge to investigate, they are presented with a
tabbed pane allowing them to view the full commit log, the files edited, the modules
involved, and the merge-tree view.

The first tab displays the full commit log (figure \ref{fig:message}). From this, a
user is able to see what they would see had they searched for the commit using git
log. This doesn't provide additional information to the other tools, but helps to
complete the functionality of \tool. The commit log provides a user with the
information about the content of the commit and who has signed-off on the commit to
ensure that it is good quality. The message for merges may contain a summary of the
commits being merged.  The information within these messages is highly variable, and
is completely dependent on the author's style. As the user moves toward the
top-level merge, the quality of these messages generally improves.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/files_view_2.png}
        \caption{Panels showing the files modified by all the commits that are part of this merge.}
        \label{fig:files}
\end{figure}

The second tab is the files tab (figure \ref{fig:files}). This tab provides information on what files have been edited, how many lines were added, and how many
lines were removed in a given commit. For non-merges, this functionality is similar to the other tools available. Our tree-based design model allows us to
extend this functionality to merges by aggregating information about all the commits that are children of the merge in the merge-tree (which the other tools are
unable to show). To find the number of lines added to a file in a merge, we take the sum of the lines added to that file in each of the children of that
merge. We do the same for calculating the number of lines removed. 
% This could be extended to use the patch information to determine when a line has been
% replaced rather than incrementing the lines added and removed.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/modules_view_2.png}
        \caption{Panel showing the modules changed by all the commits in this merge.}
        \label{fig:modules}
\end{figure}

The modules tab (figure \ref{fig:modules}) shows the modules that are contained
within the commit. Modules are not natively recognized by git, and are not going to
be present in all repositories.  In the Linux repository, authors put the module
they are working on in the one-line preview of the log-message; for example: the log \emph{gcov: add support for FCC 4.9} has updated the \emph{gcov} module
(the coverage testing tool of the kernel). We heuristically
extract the module by taking all text in the log preview of commits until the first
colon.  Modules are logical partitions of the information in the kernel. Depending
on where the author was working, modules can be general, such as ``bluetooth'' and
``wireless'', or can be quite specific for individual hardware, such as
``ath9k\_hw'' and ``wl1251''. In a few cases, the author of a commit does not
correctly follow this format and the hueristic approach fails. As with the Files panels,
non-merge commits show their corresponding information, but for merge commits we aggregate all the modules changed
in all the commits that are part of that merge.
The output of this view is shown in Figure~\ref{fig:modules}.

Finally, we have the Tree view tab. The tree view is designed for providing easy
navigation of the commits within the merge-tree that is rooted in the current merge. It also provides a clear topological view of the merge and the submerges it includes.We have experimented with various tree designs to find a design that
allows for easy navigation and visualization of both large and small trees.  We discuss this panel in the next subsection.

\subsection{Merge-Tree Views} \label{treeview_section}

The merge-tree view is what makes \tool unique to other tools that inspect the DAG of a git repository. With it, a user can inspect how commits are merged
on their way to Master. We have experimented with various types of trees:

\begin{enumerate}
\item List trees are a text-based representation of the merge-tree, and are easy to search and navigate.
\item Reingold-Tilford trees provide a clear visual represent the tree structure of the merge-tree.
\item Bubble trees organize the data hierarchically by having the parent node contain the child nodes similarly to tree maps, but
clearly showing the parent-child relationships between commits and merges.
\end{enumerate}

\evantodo{we need a reference to these trees}

To improve the visualization of the merge-tree, we first transform the merge-tree into a new tree, where each commit points to its next merge in the way to Master. However, we keep the topological order of the commits in the tree-merge as the order of the
leaves for a given merge (e.g. if merge-tree contains two commits A, B merged into C, then the simplified tree contains links B and C directly to A, but it
preserves the order of B with respect to C by showing B above C.
This allows the clustering of commits along their next merge, and avoid the space needed to
present long sequences of commits without merges in between.

The Reingold-Tilford tree (Figures \ref{fig:zoomed_tree} and \ref{fig:tree}) allows the visualization and navigation
of the entire merge-tree in an intuitive representation of the tree. This illustrates a clear notion of
root and leaves, and how to navigate in either direction. Some of the merge trees are very large (figure \ref{fig:zoomed_tree}), containing
thousands of commits and merges. While the tree is capable of producing a
visualization, it becomes much more difficult to understand. For example the merge 73287a43cc79ca06629a88d1a199cd283f42456a performed by Linus Torvalds in May
1, 2013, merged 1722 non-merge commits and 91 merges. footnote{This commit can be inspected at \url{http://li.turingmachine.org/commits/73287a43cc79ca06629a88d1a199cd283f42456a}}

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/tree_zoom.png}
        \caption{Entire Reingold-Tilford Tree of a merge. This merge by Linus (a992bf836f9c3039a16f4bd068d161c86c6c3e2c) contains 1722 includes non-merge commits
          touching 239 modules and 1525 files.}
        \label{fig:zoomed_tree}
\end{figure}

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/tree_view.png}
        \caption{Detail of the Reingold-Tilford Tree in Figure~\ref{fig:zoomed_tree}.}
        \label{fig:tree}
\end{figure}

Clicking on subtrees will cause them to collapse, only showing parts of the tree
that are of interest to the user. Furthermore, the user can zoom the tree by
scrolling the mouse wheel to see more of the topology of the tree, or to see the
details of a given part of the tree.

\dmg{do you have a figure that shows the collapsed subtrees? Try to use the same HUGE commit}

\evantodo{we need reference to bubble trees}

The bubble tree (figure \ref{fig:bubble_tree}) provides a very clear picture of
where a commit is located in the tree hierarchy. The currently selected commit or merge
is colored red, while non-selected merges use a shade of blue and commits use white.
The depth of the blue colouring is dependent on how deep in the hierarchy that node
is. The root is the lightest blue, and the contained merges are progressively
darker.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/bubble_tree.pdf}
        \caption{Bubble Tree of a large merge, showing in red (the tiny dot close to the middle) the location of a specific commit of interest.}
        \label{fig:bubble_tree}
\end{figure}

\tool was built using Nginx as the main web server, postgresql as the dbms, and Flask as the framework for generating the dynamic content.

Nginx is a high-performance, open source, scalable HTTP and mail server. It is designed with massive
scalability and multi-processing in mind. Instead of using request-driven threading,
Nginx uses an event-driven architecture, making it more scalable, proving itself as
being the web server behind large sites like Netflix, Hulu, GitHub, Tumblr, and many
other popular sites\footnote{\url{http://wiki.nginx.org/Main}}.

Postgresql is our DBMS of choice, providing fast results.

From a developers point of view, the are two major disadvantages of the DAG model of git: a) its edges point backwards, i.e. commits point to their ancestor,
not to the commit that succeeds them in their way to be integrated; and b) the DAG contains many more edges that are necessary to understand how integration
happened. We addressed these two issues with the creation of the merge-tree model of the DAG of Linux. Effectively the merge-tree view recoveres the details of each merged
branch into master.

In our experience, no other git repository reaches the level of DAG complexity that Linux has. In most git repositories  every a merge into master does not
merge other merges, and most merges bring very few new commits with them. However, even in those cases, the merge-tree will provide a valuable summary of how
commits have been integrated into master, specially since the time of integration might be very different from the time the commits were created.

The biggest challenge, however, is computing the merge-tree. Our heuristic is applicable \textbf{only} to the Linux kernel, and only since July 2009 and it is
very likely it will not work for other repositories, specially those that have fast-forward commits into Master (as the majority of repositories). One way to
address this is lack of data is by monitoring at short intervals the arrival of commits into Master (as we did in \cite{German2015}, i.e.
\emph{continuously mining} a git repo). This however, requires planning.

Regarding \tool. Because it leverages the merge-tree of the kernel, it provides mechanisms and visualizations that other tools are unable to
produce.  It gives us the ability to see how commits are being merged
into the kernel, allowing us to better understand how code is integrated into the kernel.

\begin{figure*}
        \centering
        \includegraphics[width=0.97\textwidth]{figures/042dd_tree.png}
        \caption{Merge Tree View \dmg{this figure is unreadable. Any chance you can make the text bigger?}}
        \label{fig:tree_view}
\end{figure*}

Our goal was to build a tool to enable maintainers to effectively navigate and browse
the changes performed to the kernel over the period of a release.
Achieving this goal includes removing information that does not pertain to the area
of the kernel that the user is interested in. In figure \ref{fig:dag_view} and
figure \ref{fig:tree_view} we can visually compare the results returned from gitk
and our tool for the top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6''
from kernel version 3.10. Note first that gitk has no way to show that this is a merge into Master.

The log information to this merge-commit in both of these
figures is identical, but the presentation drastically changes our ability to
comprehend what we are seeing. The primary difference is the removal of irrelevant
information. In gitk (\ref{fig:dag_view}), the visualization results include commits
and merges from other components of the kernel, while our tool (\ref{fig:tree_view})
only includes the results that are specific to the component of the kernel that we
are interested in.

This merge is relatively small, containing only four sub-merges, three of which
only contain a single commit, and the last containing two commits. With our
tool, we are able to immediately see what section of the kernel these commits and merges pertain to. The DAG view of this provides almost no explanation,
furthermore, users must work to determine where this merge ends and the next one
begins.

\tool further enhances our understanding by summarizing the files
and modules that were edited in the entire branch being merged. We are able to determine that three
files, ``palamas-regulator.c'' had two lines added and two lines removed,
``dbx500-prcmu.c'' had 12 lines added and 12 lines removed, and ``core.c'' had 5
lines added and 2 lines removed. Finally, we are able to determine that only the
``regulator'' module was modified in this merge and was modified by 5 commits.

Using \tool many advantages over the DAG model, but there are two
drawbacks. The first drawback is that a server must query the state of the master
repository since time information is not stored by git. The second drawback is the
need to pre-compute the merge-tree of the repository.

% Our core model provides advantages to the DAG model used by other tools. We separate
% the commits and merges into sub-trees and focus on a single sub-tree at a time. This
% enables us to filter out commits and merges that do not pertain to the component of
% the kernel that we are interested in, providing a clearer picture of what is
% happening in a given merge. The tree model also enables us to aggregate metadata
% easily. With the addition of the continuous polling of the repository, the model
% provides us with a mechanism of performing date-range queries with meaningful
% results.

@@
